\chapter{Existence and Uniqueness of Solutions}

In this topic, we would like to address the existence and uniqueness to the general 
first-order IVP:

\begin{equation}
    y' = f(x,y), \quad y(x_0) = y_0
\end{equation}

\begin{theorem}[Peano's Existence theorem]
    Let $R = \{(x,y) \> | \> a < x < b, c < y < d \}$ be a open rectangular 
    region containing the point $(x_0, y_0)$. If the function $f(x, y)$ 
    is continuous in $R$.

    \[
        y' = f(x,y), \quad y(x_0) = y_0
    \]

    \input{plots/chpt4-01.tex}

    in some interval $x_0 - h < x < x_0 + h$ contained in $a < x < b$.
\end{theorem}

\begin{example}
    Determine whether Peano's Existence theorem does or does not guarantee existence of a 
    solution of the initial value problem:
    \[
        xy' = y,\quad y(1) = 0
    \]
\end{example}
\begin{solution}
    The DE can be written as $y' = f(x,y)$ where $ f(x,y) = \frac{y}{x}$. 
    Observe that $f$ is continuous everywhere in the $xy$-plane except on the line 
    $x = 0$ (which is the $y$-axis). Since the initial point $(1, 0)$. Hence,
    the theorem guarantees the existence of a solution of the IVP.
\end{solution}

The next example tells us that there are first-order initial value problems that have 
more than one solutions.

\newpage

\subsection*{An IVP with more than one solution}

\begin{example}
    Verify that the function $y_1 = 0$ and $y_2 = x$ are solutions of the initial value problem

    \[
        xy' = y,\quad y(0) = 0
    \]
\end{example}

\begin{remark}
    The function $f(x,y) = \frac{y}{x}$ is continuous everywhere in the plane except
    at the points $(x, y)$ where $x=0$. Thus, Peano Existence theorem does not guarantee 
    the existence of a solution in some neighbourhood of the initial point $(0,0)$.
\end{remark}

Obviously, the next thing we would like to find out is that if an IVP does have a solution, what conditions 
could we impose on (5) to 

\begin{theorem}[Picard's Existence and Uniqueness Theorem]
    Let $R = \{(x,y) \> | \> a < x < b, c < y < d\}$ be an open rectangular region 
    containing the point $(x_0, y_0)$.

    \input{plots/chpt4-02.tex}
\end{theorem}

\begin{example}
    Determine whether Picard's theorem guarantees that the first-order IVP
    \[
        y' = y^2 + x^3, \quad y(2) = 5
    \]
    has a unique solution.
\end{example}
\begin{solution}
    Consider the following IVP

    \[
        \begin{cases}
            y' = f(x,y) = y^2 + x^3\\
            y(2) = 5
        \end{cases}
    \]

    Observe that $f$ is continuous $\forall (x,y) \in \mathbb{R}$. And since
    \[
        f_y(x, y) = \frac{\partial f}{\partial y}=2y \text{ is continuous } \forall (x,y) \in \mathbb{R}
    \]
    Thus, $f$ and $\frac{\partial f}{\partial y}$ are continuous near the initial point (2, 5). 
    By Picard's theorem, this IVP has a unique solution.
\end{solution}

\begin{example}
    Use Picard's theorem or Peano Existence theorem to discuss the existence and 
    uniqueness of the solutions of the following IVP

    \[
        y' = 3y^{2/3}, \quad y(x_0) = y_0
    \]
\end{example}

\begin{example}
    Use the Picard's existence and uniqueness theorem to prove that $y(x) = 3$ is the only 
    solution to the IVP

    \[
        y' = \frac{x(y^2 - 9)}{x^2 + 1}, \quad y(0) = 3
    \]
\end{example}

\begin{solution}
    \[
\left\{
\begin{array}{cc}
 y' = f(x,y) = \frac{x(y^2 - 9)}{x^2 + 1} & \quad \refstepcounter{equation}(1)\\
 y(x_0) = y_0 & \quad \refstepcounter{equation}(2)
\end{array}
\right.
\]

Being rational function, $f$ is continuous $\forall (x, y) \in \mathbb{R}^2$ except at the points 
where $x^2 + 1 = 0$. Since $x^2 + 1 \neq 0 \quad \forall x \in \mathbb{R}$, so $f$ is 
continuous, the same condition can be apply on $\partial f / \partial y$.

First off, we need to show that  $y(x) = 3$ is a solution to the IVP. 
By direct substitution, we substitute $y(x) = 3, y'(x) = 0$ into Eq (1).

\[
    \text{LHS of Eq } (1) = 0 = \frac{0\,(3^2 - 9)}{0^2 + 1} = \text{RHS of Eq }(1)
\]

Also, $y(x) = 3 \Rightarrow y(0) = 3 \Rightarrow y(x) = 3$ satisfies condition (2).

By Picard's theorem, $y(x) = 3$ is the only solution to the IVP.


\end{solution}